spatially-explicit model

### 2015-03-04 documentation/interactive_spexmodel.Rmd
update interactive timeseries (Shiny)

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

## developping the null model

Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.

@@

  ini_rho <- function(rho_1) {
    c(
    rho_1 = rho_1,
    rho_11 = rho_1*rho_1,
    rho_10 = rho_1*(1-rho_1)*2,
    rho_00 = (1-rho_1)*(1-rho_1),
    rho_0 = 1-rho_1
    )

  }

d_rho_1 <- function(rho, parms) { 
  with(parms,
       r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) - m * rho[1] - ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
       )  
  }

d_rho_11 <- function(rho,  parms) { 
  with(parms,
       2* (rho[1] - rho[2]) * r * (b + (f - b) * (rho[1] - rho[2])/(1- rho[1]) ) * rho[1]^( 1 + alpha) * (1 - rho[1]/(K * (1-c*(rho[1] - rho[2])/(1- rho[1])) ) ) / (1-rho[1]) - 2 * rho[2] * m  - 2 * rho[2] * ( (a + v*rho[2]/rho[1]) * rho[1]^( 1 + q) * L)/(1 +(a + v*rho[2]/rho[1]) * (h / (1- p * rho[2]/rho[1]) ) * rho[1]^( 1 + q)) 
       )  
  }

odesys_spex <- function(t, rho, parms = model_parms) {
  rho_1 <- d_rho_1(rho, parms)
  rho_11 <- d_rho_11(rho, parms)
  list(c(rho_1, rho_11, rho_1-rho_11, 1-rho_1-(rho_1-rho_11),1-rho_1  
               ) )
}

@@

As an accelerating feedback we add water runoff, which reduces growth at low cover.

@@

inputPanel(
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.1, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("alpha", label = "water runoff:",
              min = 0.0, max = 1, value = 0, step = 0.01),
  sliderInput("f", label = "local facilitation:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("c", label = "local competition:",
              min = 0.0, max = 1, value = 0.1, step = 0.01)
)

@@

@@

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 200, value = 50, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 6, step = 1),
  sliderInput("q", label = "hill coefficient:",
              min = 0.0, max = 1, value = 0, step = 0.01),
  sliderInput("v", label = "local attractivity:",
              min = 0.0, max = 1, value = 0.0, step = 0.01),
  sliderInput("p", label = "local protection:",
              min = 0.0, max = 1, value = 0.8, step = 0.01)
)

@@

@@

# running the ode-solver
ODE_spex <- ode(y = ini_rho(input$rho_1), func = odesys_spex, times = 1.1^(1:50), parms = parms)

# transfer into ouput and calculate missing rho values
ODE_spex <- as.data.frame(ODE_spex )

par(mar = c(4,4,1,3)+0.1)
plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1))
axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )

>>>>

---
title: "spatially-explicit model"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

The model includes global accelerating and decelerating feedbacks on mortality and grazing.

This panel runs all three models

- cellular automata (grey line)
- pair-approximation (black line)
- mean-field model (red line)

The pair-approximation and mean field are  solved as ordinary differential equations.

@@

defparms <- list(
    m = 0.05,
    r = 1,
    b = 0.4, 
    K = 0.9, 
    a = 0.3,
    h = 50,
    L = 5, 
    alpha = 0,
    q = 0,
    c = 0,
    f = 0,
    v = 0,
    p = 0
  )

@@

As an accelerating feedback we add water runoff, which reduces growth at low cover. 
As local feedbacks, we assume local competition, reducing the carrying capacity at high local cover, and local facilitation, compensating for poor environmental quality (a value of one restituates the maximal growth rate)

@@

inputPanel(
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = defparms$r, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = defparms$b, step = 0.0001),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = defparms$K, step = 0.01),
  sliderInput("alpha", label = "water runoff:",
              min = 0.0, max = 1, value = defparms$alpha, step = 0.01),
  sliderInput("f", label = "local facilitation:",
              min = 0.0, max = 1, value = defparms$f, step = 0.01),
  sliderInput("c", label = "local competition:",
              min = 0.0, max = 1, value = defparms$c, step = 0.01)
)

@@

Local feedbacks are local attractivity of vegetated patches, increasing the search efficiency, and local protection (associational resistance) which reduces the maximum feeding rate.

@@

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = defparms$m, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = defparms$a, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 200, value = defparms$h, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = defparms$L, step = 1),
  sliderInput("q", label = "hill coefficient:",
              min = 0.0, max = 1, value = defparms$q, step = 0.01),
  sliderInput("v", label = "local attractivity:",
              min = 0.0, max = 1, value = defparms$v, step = 0.01),
  sliderInput("p", label = "local protection:",
              min = 0.0, max = 1, value = defparms$p, step = 0.01)
)

@@

  source("../code/simfunctions.r")

@@

# running the ode-solver
ODE_spex <-  as.data.frame(ode(y = ini_rho(input$rho_1), func = odesys_spex, times = exp(seq(0,4,length = 100))-1, parms = parms))

par(mar = c(4,4,1,3)+0.1)

plot(rho_1 ~ time , data = ODE_spex , type = "l", ylim = c(0,1), xlim = c(0,50))
axis(4, at = round(tail(ODE_spex,1)$rho_1,2),cex.axis = 0.7, las = 1 )

spexmodel <- runCA(input$rho_1, parms, delta = 0.2, t_max = 60, t_min = 50, t_eval = 5, isstable = 0.001, saveeach = 1)

lines(spexmodel$rho_one ~ spexmodel$time, col = "grey50")

meanfield <-as.data.frame( ode(ini_rho(input$rho_1), func = odesys_mean, times = exp(seq(0,4,length = 100))-1, parms = parms))
lines(rho_1 ~ time , data = meanfield, col = "red3")
